---
title: "First look at Palmer Penguins"
---

# Introduction

How do bill sizes of penguins vary across species?
And what happens if we don't take species into consideration?

## Packages

We will use the **tidyverse** packages for data wrangling and visualization and the **palmerpenguins** package for the data.

```{r}
#| label: load-packages
library(tidyverse)
library(palmerpenguins)
```

## Data

The dataset we will visualize is called `penguins`.
Let's `glimpse()` at it.

```{r}
#| label: glimpse
glimpse(penguins)
```

# Penguin bills

1.  Create a scatterplot of `bill_depth` vs. `bill_length`s of `penguin`s. Overlay a line of best fit. Describe the relationship between the two variables.

```{r}
#| label: bill-depth-width
penguins |>
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Though the points show that a longer beak is typically more deep, the line of best fits indicates that there is a negative association between the two variables. This is likely because each species has distinct patterns that is affect the overall line.

2.  Create a scatterplot of `bill_depth` vs. `bill_length`s of `penguin`s, colored by `species`. Overlay a line of best fit. Describe the relationship between the two variables for each of the species.

```{r}
#| label: bill-depth-width-species
penguins |>
  ggplot(aes(
    x = bill_depth_mm,
    y = bill_length_mm,
    color = species
  )) +
  geom_point() +
  geom_smooth(method = "lm")
```

When a line of best fit is added to each spcies, it now shows the positive association for each species.

3.  Reflect on the seemingly contradictory findings from the two visualizations you've created. Which one do you believe more, and why?

It is more likely that there is a positive association between the `bill length` and `bill depth`. This is a case of *Simpsons's Paradox* in which the assocation is not fully discovered until a third variable, i.e. the species, is introduced