---
title: "Starwars characters"
---

# Introduction

You might not be a Star Wars fan, but you've probably heard of the movie franchise!

## Packages

We will use the **tidyverse** packages for data wrangling and for the data.

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
```

## Data

The dataset we will explore is called `starwars`.

# Star Wars characters

## Exercise 1

Glimpse at the starwars data frame.

```{r}
#| label: glimpse-starwars

glimpse(starwars)
```

## Exercise 2

Modify the following plot to change color the points by gender.

```{r}
#| label: scatterplot-gender

ggplot(starwars, aes(x = height, y = mass, color = gender)) +
  geom_point()
```

## Exercise 3

Modify the following plot to change the color of all points to `"pink"`.

```{r}
#| label: scatterplot-pink

ggplot(starwars, aes(x = height, y = mass, color = 'pink')) +
  geom_point()
```

## Exercise 4

Add labels for title, x and y axes, and color of points.

```{r}
#| label: scatterplot-labels

ggplot(starwars, aes(x = height, y = mass)) +
  geom_point(color = "pink")
```

## Exercise 5

Pick a single categorical variable from the data set and make a bar plot of its distribution.
Describe what you see in the plot.

Add description here.

```{r}
#| label: viz-cat

starwars |>
  filter(hair_color != "none") |>
  ggplot(aes(x = fct_infreq(hair_color))) +
  geom_bar()
```

Of the characters with hair, the majority of them have brown hair followed by black and white, respectively.

## Exercise 6

Pick two categorical variables and make a visualization to visualize the relationship between the these variables.
Along with your code and output, provide an interpretation of the visualization.

Add interpretation here.

```{r}
#| label: viz-cat-cat
starwars |> 
  filter(gender != "NA") |> 
  ggplot(aes(x = height, color =gender, fill = gender)) +
  geom_density(linewidth = 0.75, alpha = 0.5)
```
Men are generally taller than women and have a greater variation in height.

## Exercise 7

Pick three categorical variables and make a visualization to visualize the relationship between the these variables.
Along with your code and output, provide an interpretation of the visualization.

Add interpretation here.

```{r}
#| label: viz-cat-cat-cat

starwars |> ggplot(aes(x = hair_color, fill = gender)) + geom_bar() + facet_wrap(~species)
```
